Reduced plasticity and variance in physiological rates for terrestrial ectotherm populations under climate change

### 2024-02-02 ms.qmd
Methods fixed. On to results

We extracted means, standard deviations, and sample sizes for physiological rates at the two test temperatures. If there were more than two test temperatures, we chose only the test temperatures that fell within the most likely natural range of temperatures experienced by the species in question. We extracted these data from text, tables or figures of a given paper. Data were extracted from figures using the R package *metaDigitise* [@Pick2019-mg]. We also recorded the phylum, class, order, genus and species under study and the latitude and longitude of the population that was being studied. For studies that did not provide latitude and longitude for the population, we searched for similar studies by the lab group to identify where the population was likely to have been sourced or derived from when needed. If the population was derived from the wild, we recorded the nearest latitude and longitude of the population to the field collection site. If the animals had been sourced from a commercial supplier, we took the latitude and longitude of the supplier that the paper identified the animals to have originated from. When it was not possible to find latitude and longitude using these methods, we looked up the distribution of the species in question and took the latitude and longitude of the centroid of the species' distributional range.

@@

Following @Noble2022-ty we calculated a series of temperature-corrected effect sizes that compared mean physiological rates ($lnRR_{Q_{10}}$) as well as the variability in physiological rates ($lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$). These effect sizes are similar to the traditional temperature coefficient ($Q_{10}$), but with formal analytical approximations for their sampling variances. Sampling variances for effect sizes allowed us to make use of traditional meta-analytic modelling approaches.

@@

Where, $R_{1}$ and  $R_{2}$ are mean physiological rates and  $T_{1}$ and  $T_{2}$ are the temperatures that these rates are measured. Log transformation of this ratio makes the effect size normally distributed. @eq-lnq10 is essentially a temperature corrected equivalent to the log response ratio (lnRR) [@Hedges1999-de; @Lajeunesse2011-fq] when the numerator and denominator are measured at different temperatures. This allows one to compare the mean of two temperature treatments directly regardless of the temperatures that these groups have been measured. The sampling variance for @eq-lnq10 can be computed as follows (as described in @Noble2022-ty):

@@

Monthly average temperature data were extracted from the ERA5 climate model, available from the Copernicus climate data store [@Hersbach2020-hx]. For each population and species in the dataset we extracted a 30-year period (1950-2022) of either surface air temperature (0.01$^{\circ}$ resolution) for both terrestrial and freshwater taxa, or sea surface temperature for the marine taxa (at 0.25$^{\circ}$ resolution) using the *ncdf4* R package [vers. `r utils::packageVersion("ncdf4")`, @Pierce2021-lc]. We chose surface temperature because we believed that it was more likely reflects the micro-thermal environment experienced by terrestrial and freshwater ectotherms at those locations. For terrestrial species we estimated soil temperatures as this maybe more representative of microhabitat choice compared to air temperature. We fit models using both air and soil temperature and found that the results were qualitatively similar. As such, we only present results for air temperature.

@@

To understand the consequences of human-induced climate change on the potential to impact the opportunity for selection on physiogical traits we fit a model that included an interaction between acclimation type, habitat type, latitude and longitude. We assumed that any change in $lnCVR_{Q_{10}}$ across latitude and longitude could vary by habitat type (i.e., an interaction between habitat). We used non-linear tensors for latitude and longitude as any response could be complicated by local factors (e.g., altitude). Our model included random effects of species, trait, phylogeny and study. We predicted the expected change in $lnCVR_{Q_{10}}$ for each wild population in our dataset at the specific populations latitude and longitude. To do this, we first converted the predicted $lnCVR_{Q_{10}}$ to a 1$^{\circ}$C change as opoosed to 10$^{\circ}$C as follows:

@@

We explored the possibility for publication bias graphically, using funnel plots, and more formally by including in our meta-regression models sampling variance (or sampling standard error) [@Nakagawa2022-gy]. Funnel plot asymmetry may suggest a form of publication bias called the 'file-drawer' effect whereby low-powered studies are less likely to be published. To test whether sampling variance covaried with effect size we included it in a multi-level meta-regression model that acounted for all the random effects (study, species, trait) and fixed effects (acclimation time, type of effect, habitat, trait category and the interaction between habitat type and trait category).

>>>>

We extracted means, standard deviations, and sample sizes for physiological rates at the two test temperatures. If there were more than two test temperatures, we chose only the test temperatures that fell within the most likely natural range of temperatures experienced by the species in question. We extracted these data from text, tables or figures of a given paper. Data were extracted from figures using the R package *metaDigitise* [@Pick2019-mg]. We also recorded the phylum, class, order, genus and species, and the latitude and longitude from where the experimental animals were sourced. For studies that did not provide latitude and longitude for the population, we searched for similar studies by the same lab group to identify where the population was likely to have been sourced. If the experimental animals were derived from the wild, we recorded the nearest latitude and longitude of the field collection site. If the animals were sourced from a commercial supplier, we took the latitude and longitude of the supplier. When it was not possible to find latitude and longitude using these methods, we looked up the distribution of the species in question and took the latitude and longitude of the centroid of the species' distributional range.

@@

Following @Noble2022-ty we calculated a series of temperature-corrected effect sizes that compared mean physiological rates ($lnRR_{Q_{10}}$) as well as the variability in physiological rates ($lnVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$). These effect sizes are similar to the traditional temperature coefficient ($Q_{10}$), but with formal analytical approximations of their sampling variances. Sampling variances for effect sizes allowed us to make use of traditional meta-analytic modelling approaches.

@@

Where, $R_{1}$ and  $R_{2}$ are mean physiological rates and  $T_{1}$ and  $T_{2}$ are the temperatures at which these rates are measured. Log transformation of this ratio makes the effect size normally distributed. @eq-lnq10 is essentially a temperature corrected equivalent to the log response ratio (lnRR) [@Hedges1999-de; @Lajeunesse2011-fq] when the numerator and denominator are measured at different temperatures. This allows comparisons of the means from two temperature treatments directly regardless of the absolute measurement temperatures. The sampling variance for @eq-lnq10 can be computed as follows (as described in @Noble2022-ty):

@@

Monthly average temperature data were extracted from the ERA5 climate model, available from the Copernicus climate data store [@Hersbach2020-hx]. For each population and species in the dataset we extracted a 30-year period (1950-2022) of either surface air temperature (0.01$^{\circ}$ resolution) for both terrestrial and freshwater taxa, or sea surface temperature for the marine taxa (at 0.25$^{\circ}$ resolution) using the *ncdf4* R package [vers. `r utils::packageVersion("ncdf4")`, @Pierce2021-lc]. We chose surface temperature because we believed that it was more likely to reflect the micro-thermal environment experienced by terrestrial and freshwater ectotherms at those locations. For terrestrial species we estimated soil temperatures as this maybe more representative of microhabitat choice compared to air temperature. We fit models using both air and soil temperature and found that the results were qualitatively similar. We therefore only present results for air temperature.

@@

To understand the consequences of human-induced climate change on the potential to impact the opportunity for selection on physiological traits we fit a model that included an interaction between acclimation type, habitat type, latitude and longitude. We assumed that any change in $lnCVR_{Q_{10}}$ across latitude and longitude could vary by habitat type (i.e., an interaction between habitat). We used non-linear tensors for latitude and longitude as any response could be complicated by local factors (e.g., altitude). Our model included random effects of species, trait, phylogeny and study. We predicted the expected change in $lnCVR_{Q_{10}}$ for each wild population in our dataset at the specific populations latitude and longitude. To do this, we first converted the predicted $lnCVR_{Q_{10}}$ to a 1$^{\circ}$C change as opoosed to 10$^{\circ}$C as follows:

@@

We explored the possibility for publication bias graphically, using funnel plots, and more formally by including in our meta-regression models sampling variance (or sampling standard error) [@Nakagawa2022-gy]. Funnel plot asymmetry may suggest a form of publication bias called the 'file-drawer' effect whereby low-powered studies are less likely to be published. To test whether sampling variance covaried with effect size we included it in a multi-level meta-regression model that accounted for all the random effects (study, species, trait) and fixed effects (acclimation time, type of effect, habitat, trait category and the interaction between habitat type and trait category).
